|

### 2023-01-27 paper.tex
autogit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2023-01-11}

@@

\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage  
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

DAE-LSTMs are recurrent neural networks (RNNs) that can be trained in an
unsupervised learning environment to minimize a reconstruction error for
encoding input data into a compressed representation and subsequently
decoding the compressed representation back to the input dimension. In
DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions
selected from a parent population with the aim to capture their
promising qualities. The resulting model is then used to sample new
offspring solutions by propagating partially mutated solutions from the
parent population through the DAE-LSTMS
{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}. In previous work
DAE-GP has been shown to outperform GP for both a generalized version of
the royal tree problem
{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a
real-world symbolic regression problem
{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.

The DAE-GP algorithm first described by
{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs
for each generation \(g\) of the search from scratch.
{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and
{[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the
incorporation of a pre-training strategy into the evolutionary search as
a possible way of improving the model quality of DAE-GP. The key idea is
to pre-train an initial DAE-LSTM on a large population of candidate
solutions and to use the pre-trained parameters of this initial model in
each generation of the search as starting parameters for the current
generation DAE-LSTM. This thesis studies the influence of using
pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially
looking at the influence in generalization behavior and the overall
quality of solutions found by DAE-GP. The aim of this study is to answer
the question if a pre-training strategy can help to improve
generalization in DAE-GP and if so, analyze how this improved
generalization ability benefits the overall performance of DAE-GP.

@@

DAE was first introduced by
{[}\protect\hyperlink{ref-10.1145ux2f1390156.1390294}{14}{]} as an
improved AE with the ability to learn new representations of data that
is especially robust to partially corrupted input data. DAE modifies the
AE by using partially corrupted input data for the AE and training it to
reconstruct the uncorrupted, original version of the input data.

@@

LSTMs are a variant of recurrent neural networks first introduced by
{[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned
information over en extended time period while avoiding the problem of
vanishing and/or exploding gradients. Since DAE-GP encodes candidate
solutions as linear strings in prefix notation, the DAE in DAE-GP uses
LSTMs for both encoding and decoding where the total amount of time
steps \(T\) is equal to sum of the length of the input solution and the
output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.

@@

Pre-Training describes the concept of initially training a neural
network on a large dataset before using it to solve a more specific
task. It is a commonly used strategy in deep architectures that has been
shown to improve both the optimization process itself as well as the
generalization behavior if compared to the standard approach of using
randomly initialized parameters
{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.

@@

Another important finding by
{[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive
effect of pre-training described above is dependent on both the depth of
the network as well as the size of layers - while increasingly larger
networks benefit increasingly more from pre-training the final
performance of small architectures tends to be worse with pre-training
than with randomized weight initialization. The authors find evidence
that pre-training DAE is especially usefull for optimizing the
parameters of the lower layers of the network.

@@

The DAE-GP algorithm, first described by
{[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a
flowchart in figure §. After setting the generation counter \(g\) to
\(0\) and creating the initial population of solutions \(P_0\), DAE-GP
enters the main loop that checks if a termination criteria is satisfied
(i.e.~maximum number of generations or fitness evaluations is reached).
For each generation \(g\) the fitness of all new individuals in \(P_g\)
is evaluated. During model building, DAE-GP selects a subset \(X_g\) of
the current population that is used as training data for the DAE-LSTM
model \(M_g\) which is trained to learn the properties of \(X_g\). The
trained model \(M_g\) is then used for model sampling, partially mutated
solutions from \(P_G\) are propagated through \(M_g\) to create a new
population \(P_{g+1}\). This process is repeated until the termination
criteria is met and DAE-GP returns the highest fitness solutions.

\begin{figure}
\centering
\includegraphics{./img/flowcharts/dae-gp.png}
\caption{Flowchart - Regular DAE-GP}
\end{figure}

@@

\caption{\label{tab:airfoil_dataset_description}Airfoil - Dataset Description}
\centering
\begin{tabular}[t]{l|l|l|l}
\hline
\textbf{Type} & \textbf{Name} & \textbf{Description} & \textbf{Unit}\\
\hline
input & x1 & Frequency & Hertz\\
\hline
input & x2 & Angle of attack & Degree\\
\hline
input & x3 & Chord length & meters\\
\hline
input & x4 & Free-stream velocity & meters/second\\
\hline
input & x5 & Suction side displacement thickness & meters\\
\hline
output & y & Scaled sound pressure level & decibels\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:airfoil_function_set}Airfoil - Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{function.} & \textbf{arity}\\
\hline
addition & 2\\
\hline
subtraction & 2\\
\hline
multiplication & 2\\
\hline
analytic\_quotient & 2\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:airfoil_fullRun_2hl_maxIndSize_params}Airfoil - Parameter for full Run}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{parameter} & \textbf{value}\\
\hline
populationSize & 500\\
\hline
generations & 30\\
\hline
fitness & RMSE\\
\hline
TrainingMode & Convergence\\
\hline
SamplingSteps & 1\\
\hline
hiddenLayers & 2\\
\hline
Selection & Binary Tournament Selection\\
\hline
Pre-Training PopulationSize(Training/Validation) & 10000/100000\\
\hline
Pre-Training TrainingMode & Early Stopping\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that
DAE-GP finds higher quality that are also smaller in size than those
found by average GP for a given number of 10.000 fitness evaluations, I
also studied the differences in average solution size between
pre-trained and regular DAE-GP.

\begin{figure}
\centering
\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
\caption{Average Solution Size over 30 Generations - Airfoil}
\end{figure}

@@

\caption{\label{tab:full_run_realWorldSymReg_problems}Overview - Real World Symbolic Regression Benchmark Problems}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
\hline
Airfoil & 1503 & 5\\
\hline
Boston\_Housing & 506 & 13\\
\hline
Energy\_Cooling & 768 & 8\\
\hline
Concrete & 1030 & 8\\
\hline
\end{tabular}
\end{table}

Table X summarizes the median best fitness achieved after running each
algorithm for 10 runs using a population size of 500 individuals for 30
generations.\footnote{P-Values are marked with asteriks (1,2,3) for
  (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1910}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1685}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0787}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1461}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1124}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1573}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Hidden-Layers
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Set
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
DAE-GP
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pre-Trained
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
P-Value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Cliffs-Delta
\end{minipage} \\
\midrule()
\endhead
Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
& 1 & Test & \textbf{33.7271} & 35.612 & 0.1400 & 0.4 \\
Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
& 2 & Test & 17.5299 & \textbf{14.7603} & 0.5706 & -0.16 \\
Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.2894 &
-0.29 \\
& 2 & Test & 8.0501 & \textbf{7.9692} & 0.7051 & -0.11 \\
Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.0188** &
0.63 \\
& 2 & Test & \textbf{4.67} & 4.9196 & 0.2886 & 0.29 \\
Concrete & 2 & Train & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
& 2 & Test & \textbf{15.7978} & 16.7196 & 0.6667 & 1.0 \\
\bottomrule()
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2152}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1519}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Hid.Layers
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
DAE-GP
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pre-Trained
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
P-Value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Cliffs-Delta
\end{minipage} \\
\midrule()
\endhead
Airfoil & 1 & \textbf{33.3674} & 35.3454 & 0.0171** & 0.64 \\
Airfoil & 2 & 17.4762 & \textbf{14.7233} & 0.3073 & -0.28 \\
Boston\_Housing & 2 & 8.1922 & \textbf{8.0763} & 0.2894 & -0.29 \\
Energy(Cooling) & 2 & \textbf{4.5271} & 4.7782 & 0.0188** & 0.63 \\
Concrete & 2 & \textbf{17.1075} & 17.528 & 1.0000 & 0.0 \\
\bottomrule()
\end{longtable}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2023-01-27}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\setcounter{page}{1}
\tableofcontents
\thispagestyle{empty}
\newpage
\listoftables
\listoffigures
\thispagestyle{empty}
\newpage

@@

DAE-LSTMs are artificial neural networks that can be trained in an unsupervised learning environment to minimize a reconstruction error for encoding input data into a compressed representation and subsequently decoding the compressed representation back to the input dimension.
In DAE-GP, DAE-LSTMs are trained with a subset of high-fitness solutions selected from a parent population with the aim to capture their promising qualities.
The resulting model is then used to sample new offspring solutions by propagating partially mutated solutions from the parent population through the DAE-LSTMS {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.
In previous work DAE-GP has been shown to outperform GP for both a generalized version of the royal tree problem {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} as well as for a real-world symbolic regression problem {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}.

The DAE-GP algorithm first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]} trains a DAE-LSTMs for each generation \(g\) of the search from scratch.
{[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} and {[}\protect\hyperlink{ref-daegp_explore_exploit}{16}{]} suggests the incorporation of a pre-training strategy into the evolutionary search as a possible way of improving the overall performance of the DAE-GP algorithm.
The key idea is to pre-train an initial DAE-LSTM on a large population of candidate solutions and to use the pre-trained parameters of this initial model in each generation of the search as starting parameters for the current generation DAE-LSTM.
This thesis studies the influence of using pre-trained DAE-LSTMs in DAE-GP for symbolic regression, especially looking at the influence on overall quality of solutions found by DAE-GP and effects on runtime.
The aim of this study is to answer the question if a pre-training strategy can be used to improve DAE-GP and to study both positive and negative effects on overall performance.

@@

DAE was first introduced by {[}\protect\hyperlink{ref-dae_orig2008}{14}{]} as an improved AE with the ability to learn new representations of data that is especially robust to partially corrupted input data. DAE modifies the AE by using partially corrupted input data for the AE and training it to reconstruct the uncorrupted, original version of the input data.

@@

LSTMs are a variant of artificial neural networks first introduced by {[}\protect\hyperlink{ref-lstm_orig}{7}{]} that can store learned information over en extended time period while avoiding the problem of vanishing and/or exploding gradients. Since DAE-GP encodes candidate solutions as linear strings in prefix notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding where the total amount of time steps \(T\) is equal to sum of the length of the input solution and the output solution {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}.

@@

Pre-Training describes the concept of initially training an artificial neural network on a large dataset before using it to solve a more specific task.
It is a commonly used strategy in deep architectures that has been shown to improve both the optimization process itself as well as the generalization behavior if compared to the standard approach of using randomly initialized parameters {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]}.

@@

Another important finding by {[}\protect\hyperlink{ref-pmlr-v5-erhan09a}{3}{]} is that the positive effect of pre-training is dependent on both the depth of the network as well as the size of layers - while increasingly larger networks benefit increasingly more from pre-training the final performance of small architectures tends to be worse with pre-training than with randomized weight initialization.
The authors find evidence that pre-training DAE is especially usefull for optimizing the parameters of the lower layers of the network.

@@

The DAE-GP algorithm, first described by {[}\protect\hyperlink{ref-dae-gp_2020_rtree}{18}{]}, is summarized as a flowchart in \ref{fig:daegp-flowchart}.
After setting the generation counter \(g\) to \(0\) and creating the initial population of solutions \(P_0\), DAE-GP enters the main loop that checks if a termination criteria is satisfied (i.e.~maximum number of generations or fitness evaluations is reached).
For each generation \(g\) the fitness of all new individuals in \(P_g\) is evaluated. During model building, DAE-GP selects a subset \(X_g\) of the current population that is used as training data for the DAE-LSTM model \(M_g\) which is trained to learn the properties of \(X_g\).
The trained model \(M_g\) is then used for model sampling, partially mutated solutions from \(P_G\) are propagated through \(M_g\) to create a new population \(P_{g+1}\).
This process is repeated until the termination criteria is met and DAE-GP returns the highest fitness solutions.

\begin{figure}[c]

{\centering \includegraphics[width=0.5\linewidth,height=0.9\textheight]{./img/flowcharts/dae-gp}

}

\caption{Regular DAE-GP Flowchart}\label{fig:daegp-flowchart}
\end{figure}

@@

\caption{\label{tab:airfoil-dataset-description}Airfoil - Dataset Description}
\centering
\begin{tabular}[t]{l|l|l|l}
\hline
\textbf{Type} & \textbf{Name} & \textbf{Description} & \textbf{Unit}\\
\hline
input & x1 & Frequency & Hertz\\
\hline
input & x2 & Angle of attack & Degree\\
\hline
input & x3 & Chord length & meters\\
\hline
input & x4 & Free-stream velocity & meters/second\\
\hline
input & x5 & Suction side displacement thickness & meters\\
\hline
output & y & Scaled sound pressure level & decibels\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:airfoil-function-set}Airfoil - Function Set}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{operator} & \textbf{arity}\\
\hline
addition & 2\\
\hline
subtraction & 2\\
\hline
multiplication & 2\\
\hline
analytic\_quotient & 2\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:airfoil-fullRun-2hl-maxIndSize-params}Airfoil - Parameter for full Run}
\centering
\begin{tabular}[t]{l|l}
\hline
\textbf{Hyperparameter} & \textbf{Value}\\
\hline
popSize & 500\\
\hline
generations & 30\\
\hline
fitness & RMSE\\
\hline
TrainingMode & Convergence\\
\hline
SamplingSteps & 2\\
\hline
hiddenLayers & 2\\
\hline
Selection & Binary Tournament Selection\\
\hline
corruptionTechnique & Levenshtein Edit\\
\hline
edit\_prob(training/sampling) & 0.05/0.95\\
\hline
functionSet & \{+ ; - ; * ; aq\}\\
\hline
Pre-Training PopulationSize & 10000\\
\hline
Pre-Training Train/Test Split & 50\%\\
\hline
Pre-Training TrainingMode & Early Stopping\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png}
Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.

\begin{figure}
\centering
\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
\caption{Average Solution Size over 30 Generations - Airfoil}
\end{figure}

@@

\caption{\label{tab:full-run-realWorldSymReg-problems}Overview - Real World Symbolic Regression Benchmark Problems}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Problem} & \textbf{Observations} & \textbf{Features}\\
\hline
Airfoil & 1503 & 5\\
\hline
Boston\_Housing & 506 & 13\\
\hline
Energy\_Cooling & 768 & 8\\
\hline
Concrete & 1030 & 8\\
\hline
\end{tabular}
\end{table}

Table X summarizes the median best fitness achieved after running each algorithm for 10 runs using a population size of 500 individuals for 30 generations.\footnote{P-Values are marked with asteriks (1,2,3) for (\(\alpha\))-levels of \(0.1\), \(0.05\) and \(0.01\)}. The median best fitness over all 30 generations are visualized in figure X.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1932}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1705}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0795}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1477}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1023}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1591}}@{}}
\caption{\label{tab:full-run-realWorldSymReg-fitness}Median Best Fitness - Symbolic Regression}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Hidden-Layers
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Set
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
DAE-GP
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pre-Trained
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
P-Value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Cliffs-Delta
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Problem
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Hidden-Layers
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Set
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
DAE-GP
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pre-Trained
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
P-Value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Cliffs-Delta
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Airfoil & 1 & Train & \textbf{33.3674} & 35.3454 & 0.02** & 0.64 \\
& 1 & Test & \textbf{33.7271} & 35.612 & 0.14 & 0.40 \\
Airfoil & 2 & Train & 17.4762 & \textbf{14.7233} & 0.31 & -0.28 \\
& 2 & Test & 17.5299 & \textbf{14.7603} & 0.57 & -0.16 \\
Boston\_Housing & 2 & Train & 8.1922 & \textbf{8.0763} & 0.29 & -0.29 \\
& 2 & Test & 8.0501 & \textbf{7.9692} & 0.71 & -0.11 \\
Energy(Cooling) & 2 & Train & \textbf{4.5271} & 4.7782 & 0.02** & 0.63 \\
& 2 & Test & \textbf{4.67} & 4.9196 & 0.29 & 0.29 \\
Concrete & 2 & Train & 17.0052 & \textbf{16.9085} & 0.97 & 0.02 \\
& 2 & Test & \textbf{16.5845} & 16.9298 & 0.52 & 0.18 \\
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_fitness_byGens.png}
\caption{Median Best Fitness - Symbolic Regression}
\end{figure}

@@

Next I studied the size of solutions based on using a pre-training strategy. Figure X shows the median size of the best solutions aggregated of all benchmark problems for symbolic regression

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=0.8\textheight]{~/masterThesis/master_thesis/img/symbolicRegressionSummarized/median_bestSol_size_byGens.png}
\caption{Median Solution Size - Symbolic Regression}
\end{figure}

Table X summarizes the results of the statistical tests.
